---
title: "Cogsci Plots"
output: word_document
date: "2026 Jan"
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

library('groundhog')
groundhog.library(
  c(
    'papaja',
    'tidyverse'
  ), "2025-06-01"
)

r_refs("r-references.bib")

knitr::opts_chunk$set(warning=F,echo=F,message=F,cache=T)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```



```{r loadformat, include = FALSE, cache=F}
library(tidyverse)
library(stringr)
library(effsize)

two_parts_df_exp1 <-read.csv("../data/experiment1/two_parts_df_exp1.csv")

two_parts_df_exp2 <- read.csv("../data/experiment2/two_parts_df_exp2.csv")

two_parts_df_exp3 <- read.csv("../data/experiment3/two_parts_df_exp3.csv")


p1_colour <- '#69b3a2'
p2_colour <- '#404080'

```

```{r pstone, cache=F}
library(broom)

exp1_pstone_slopes_p1 <- two_parts_df_exp1 %>%
  group_by(subj_id, mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  do(tidy(glm(chose_onlystones ~ trial_num, family = binomial, data = .))) %>%
  filter(term == "trial_num") %>%
  select(subj_id, slope_p1 = estimate)

exp1_pstone_slopes_p2 <- two_parts_df_exp1 %>%
  group_by(subj_id, mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  do(tidy(glm(chose_onlystones_p2 ~ trial_num, family = binomial, data = .))) %>%
  filter(term == "trial_num") %>%
  select(subj_id, slope_p2 = estimate)

exp2_pstone_slopes_p1 <- two_parts_df_exp2 %>%
  group_by(subj_id, mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  do(tidy(glm(chose_onlystones ~ trial_num, family = binomial, data = .))) %>%
  filter(term == "trial_num") %>%
  select(subj_id, slope_p1 = estimate)

exp2_pstone_slopes_p2 <- two_parts_df_exp2 %>%
  group_by(subj_id, mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  do(tidy(glm(chose_onlystones_p2 ~ trial_num, family = binomial, data = .))) %>%
  filter(term == "trial_num") %>%
  select(subj_id, slope_p2 = estimate)

exp1_diff_slope <- exp1_pstone_slopes_p1 %>%
  inner_join(exp1_pstone_slopes_p2, by = "subj_id") %>%
  mutate(diff = slope_p1 - slope_p2)

exp2_diff_slope <- exp2_pstone_slopes_p1 %>%
  inner_join(exp2_pstone_slopes_p2, by = "subj_id") %>%
  mutate(diff = slope_p1 - slope_p2)

wilcox.test(exp1_diff_slope$diff)

wilcox.test(exp2_diff_slope$diff)


```

Experiment 1, part 1, pstone slope: `r apa_print(wilcox.test(exp1_pstone_slopes_p1$slope_p1))$statistic`

Experiment 1, part 2, pstone slope: `r apa_print(wilcox.test(exp1_pstone_slopes_p2$slope_p2))$statistic`

Experiment 2, part 1, pstone slope: `r apa_print(wilcox.test(exp2_pstone_slopes_p1$slope_p1))$statistic`

Experiment 2, part 2, pstone slope: `r apa_print(wilcox.test(exp2_pstone_slopes_p2$slope_p2))$statistic`

Experiment 1, part 1 - part 2, pstone slope : `r apa_print(wilcox.test(exp1_diff_slope$diff))$statistic`

Experiment2, part 1 - part 2, pstone slope: `r apa_print(wilcox.test(exp2_diff_slope$diff))$statistic`

Exp. 1 vs Exp. 2, (part 1 - part 2), pstone slope: `r apa_print(wilcox.test(exp2_diff_slope$diff, exp1_diff_slope$diff))$statistic`


```{r no_diff_reconstruction_accuracy, cache=F}

diff_acc_exp1 <- two_parts_df_exp1 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    acc = mean(is_consistent)
  ) %>%
  tidyr::pivot_wider(
    names_from = chose_gems,
    values_from = acc,
    names_prefix = "chose_gems_"
  ) %>%
  mutate(
    diff_acc = chose_gems_TRUE - chose_gems_FALSE
  ) 

diff_acc_exp2 <- two_parts_df_exp2 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    acc = mean(is_consistent)
  ) %>%
  tidyr::pivot_wider(
    names_from = chose_gems,
    values_from = acc,
    names_prefix = "chose_gems_"
  ) %>%
  mutate(
    diff_acc = chose_gems_TRUE - chose_gems_FALSE
  ) 
```

There is no difference in the effect of reconstruction accuracy being boosted by outcomes in part 1 across experiment 1 and 2 (`r apa_print(t.test(diff_acc_exp1$diff_acc, diff_acc_exp2$diff_acc))$statistic`)

### phase 1 accuracy

```{r t_test, cache=F}

library(tidyverse)
library(broom)

make_bins <- function(max_trial, bin_width = 5) {
  breaks <- seq(0.5, max_trial + 0.5, by = bin_width)
  starts <- seq(1, max_trial, by = bin_width)
  ends   <- pmin(starts + bin_width - 1, max_trial)
  labels <- paste0(starts, "–", ends)
  list(breaks = breaks, labels = labels)
}

run_bin_ttests <- function(df, max_trial, bin_width = 5) {
  bins <- make_bins(max_trial = max_trial, bin_width = bin_width)
  breaks <- bins$breaks
  bin_labels <- bins$labels

  df_long <- df %>%
    pivot_longer(
      cols = c(chose_onlystones, chose_onlystones_p2),
      names_to = "phase_raw",
      values_to = "chose_onlystones"
    ) %>%
    mutate(
      phase = case_when(
        phase_raw == "chose_onlystones"    ~ "part1",
        phase_raw == "chose_onlystones_p2" ~ "part2",
        TRUE ~ NA_character_
      ),
      phase = factor(phase, levels = c("part1", "part2")),
      bin = cut(
        trial_num,
        breaks = breaks,
        labels = bin_labels,
        include.lowest = TRUE,
        right = TRUE
      ),
      bin = factor(bin, levels = bin_labels)
    ) %>%
    filter(!is.na(phase), !is.na(bin))

  by_subj_bias <- df_long %>%
    group_by(subj_id, phase, bin) %>%
    summarise(bias = mean(chose_onlystones, na.rm = TRUE), .groups = "drop")

  bin_tests <- by_subj_bias %>%
    pivot_wider(names_from = phase, values_from = bias) %>%
    group_by(bin) %>%
    group_modify(~ {
      dat <- .x

      dat_pair <- dat %>% filter(!is.na(part1) & !is.na(part2))

      n_subj <- nrow(dat_pair)
      mean1 <- if (n_subj > 0) mean(dat_pair$part1, na.rm = TRUE) else NA_real_
      mean2 <- if (n_subj > 0) mean(dat_pair$part2, na.rm = TRUE) else NA_real_
      diff_mean <- if (n_subj > 0) mean(dat_pair$part1 - dat_pair$part2, na.rm = TRUE) else NA_real_

      if (n_subj >= 2) {
        t_res <- t.test(dat_pair$part1, dat_pair$part2, paired = TRUE)
        t_tidy <- broom::tidy(t_res)
        tibble(
          n_subjects = n_subj,
          mean_part1 = mean1,
          mean_part2 = mean2,
          mean_diff = diff_mean,             
          t = t_tidy$statistic,
          df = t_tidy$parameter,
          p.value = t_tidy$p.value,
          conf.low = t_tidy$conf.low,
          conf.high = t_tidy$conf.high
        )
      } else {
        tibble(
          n_subjects = n_subj,
          mean_part1 = mean1,
          mean_part2 = mean2,
          mean_diff = diff_mean,
          t = NA_real_,
          df = NA_real_,
          p.value = NA_real_,
          conf.low = NA_real_,
          conf.high = NA_real_
        )
      }
    }, .keep = TRUE) %>%
    ungroup() %>%
    arrange(bin)

  bin_tests <- bin_tests %>%
    mutate(bin = factor(as.character(bin), levels = bin_labels))

  return(bin_tests)
}

results_exp1 <- run_bin_ttests(
  df = two_parts_df_exp1 %>% filter(!is_catch_trial),
  max_trial = 60,
  bin_width = 5
)

results_exp2 <- run_bin_ttests(
  df = two_parts_df_exp2 %>% filter(!is_catch_trial),
  max_trial = 50,
  bin_width = 5
)

results_exp3 <- run_bin_ttests(
  df = two_parts_df_exp3 %>% filter(!is_catch_trial),
  max_trial = 50,
  bin_width = 5
)

results_exp1_explicit <- run_bin_ttests(
  df = two_parts_df_exp1 %>% filter(!is_catch_trial, mentioned_rule),
  max_trial = 60,
  bin_width = 5
)

results_exp2_explicit <- run_bin_ttests(
  df = two_parts_df_exp2 %>% filter(!is_catch_trial, mentioned_rule),
  max_trial = 50,
  bin_width = 5
)

results_exp3_explicit <- run_bin_ttests(
  df = two_parts_df_exp3 %>% filter(!is_catch_trial, mentioned_rule),
  max_trial = 50,
  bin_width = 5
)



```

```{r phase_1_and_2_accuracy,  cache=F, fig.width=4, fig.height=6}

library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(ggtext) 

make_bins <- function(max_trial, bin_width = 5) {
  breaks <- seq(0.5, max_trial + 0.5, by = bin_width)
  starts <- seq(1, max_trial, by = bin_width)
  ends   <- pmin(starts + bin_width - 1, max_trial)
  labels <- paste0(starts, "–", ends)
  list(breaks = breaks, labels = labels)
}

pd <- position_dodge(width = 0.2)
phase_pal <- c("part1" = p1_colour, "part2" = p2_colour)  
vline_geom <- function(n_bins) seq(1.5, n_bins - 0.5, by = 1)

make_phase_plot_dynamic <- function(df,
                                    max_trial,
                                    plot_subtitle = NULL,
                                    plot_title = NULL,
                                    bin_width = 5,
                                    ylim = c(0.15,0.37)) {
  bins <- make_bins(max_trial = max_trial, bin_width = bin_width)
  breaks <- bins$breaks
  bin_labels <- bins$labels
  n_bins <- length(bin_labels)
  vline_x <- vline_geom(n_bins)
  
  df_long <- df %>%
    pivot_longer(
      cols = c(chose_onlystones, chose_onlystones_p2),
      names_to = "phase_raw",
      values_to = "chose_onlystones"
    ) %>%
    mutate(
      phase = case_when(
        phase_raw == "chose_onlystones"    ~ "part1",
        phase_raw == "chose_onlystones_p2" ~ "part2",
        TRUE ~ NA_character_
      ),
      phase = factor(phase, levels = c("part1", "part2")),
      bin = cut(
        trial_num,
        breaks = breaks,
        labels = bin_labels,
        include.lowest = TRUE,
        right = TRUE
      ),
      bin = factor(bin, levels = bin_labels)
    ) %>%
      filter(!is.na(phase))  
  
  by_subj_bias <- df_long %>%
    group_by(subj_id, phase, bin) %>%
    summarise(bias = mean(chose_onlystones, na.rm = TRUE), .groups = "drop")
  
  group_summary_sub <- by_subj_bias %>%
    group_by(phase, bin) %>%
    summarise(
      m  = mean(bias, na.rm = TRUE),
      se = sd(bias, na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )
  
  p <- ggplot(group_summary_sub,
              aes(x = bin, y = m,
                  color = phase,
                  shape = phase,
                  group = phase)) +
    geom_hline(yintercept = 1/3, linewidth = 0.3) +
    geom_line(linewidth = 0.7, position = pd) +
    geom_point(size = 2, position = pd) +
    geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0.2, position = pd, linetype = "solid") +
   scale_color_manual(
  values = phase_pal,
  name = NULL,
  labels = c(
    "part1" = "Collect the gems!",
    "part2" = "Repeat your guesses!"
  )
) +
scale_shape_discrete(
  name = NULL,
  labels = c(
    "part1" = "Collect the gems!",
    "part2" = "Repeat your guesses!"
  )
) +
scale_linetype_discrete(
  name = NULL,
  labels = c(
    "part1" = "Collect the gems!",
    "part2" = "Repeat your guesses!"
  )
) +
    scale_y_continuous(limits=ylim)+
    labs(x = "Trial Bin", y = "P(*stones only*)", title = plot_title, subtitle = plot_subtitle) + theme_minimal(base_size = 12) +
  theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
    legend.position = "none",
    axis.title.y = element_markdown(size = 14),
    axis.title.x = element_markdown(size = 14),
    axis.text.x = element_text(size = 12, angle=45, hjust = 1),
    axis.text.y = element_text(size = 12),
  ) 
  
  return(p)
}

plot_exp1 <- make_phase_plot_dynamic(
  two_parts_df_exp1 %>% filter(!is_catch_trial),
  max_trial = 60,
  plot_title = NULL,
  plot_subtitle = "Exp. 1" 
)

plot_exp2 <- make_phase_plot_dynamic(
  two_parts_df_exp2 %>% filter(!is_catch_trial),
  max_trial = 50,
  plot_title = NULL,
  plot_subtitle = "Exp. 2"
)

plot_exp3 <- make_phase_plot_dynamic(
  two_parts_df_exp3 %>% filter(!is_catch_trial),
  max_trial = 50,
  plot_title = NULL,
  plot_subtitle = "Exp. 3"
)


combined <- plot_exp1 / plot_exp2 / plot_exp3 + 
  plot_layout(guides = "collect", ncol = 1) +
  plot_annotation(
    theme = theme(plot.title = element_text(size = 16, face = "bold"),
                  plot.caption = element_text(size = 8))
  )


print(combined)

```

```{r t_plot, cache=F, fig.width=4, fig.height=6}

make_sig_layer <- function(ttest_df, y = 0.15, pad = 0.35, colour = "red", linewidth = 0.8) {
  sig_bins <- ttest_df %>%
    filter(!is.na(p.value), p.value < 0.05) %>%
    mutate(pos = as.integer(bin)) %>%
    arrange(pos)

  if (nrow(sig_bins) == 0) {
    return(geom_blank())
  }

  runs <- split(sig_bins, cumsum(c(1, diff(sig_bins$pos) != 1)))
  seg_df <- map_dfr(runs, function(df_run) {
    start_pos <- min(df_run$pos)
    end_pos   <- max(df_run$pos)
    tibble(
      x = start_pos - pad,
      xend = end_pos + pad,
      y = y,
      yend = y
    )
  })

  geom_segment(
    data = seg_df,
    aes(x = x, xend = xend, y = y, yend = yend),
    inherit.aes = FALSE,
    color = colour,
    linewidth = linewidth,
    lineend = "round"
  )
}


learn_plot_exp1 <- make_phase_plot_dynamic(
  two_parts_df_exp1 %>% filter(!is_catch_trial),
  max_trial = 60,
  plot_title = NULL,
  plot_subtitle = ""
) 
#+
#  make_sig_layer(results_exp1, y = 0.15)

learn_plot_exp2 <- make_phase_plot_dynamic(
  two_parts_df_exp2 %>% filter(!is_catch_trial),
  max_trial = 50,
  plot_title = NULL,
  plot_subtitle = ""
) 
#+
#  make_sig_layer(results_exp2, y = 0.15)

learn_plot_exp3 <- make_phase_plot_dynamic(
  two_parts_df_exp3 %>% filter(!is_catch_trial),
  max_trial = 50,
  plot_title = NULL,
  plot_subtitle = ""
) +
  make_sig_layer(results_exp3, y = 0.15)
combined <- learn_plot_exp1 / learn_plot_exp2 / learn_plot_exp3 + 
  plot_layout(guides = "collect", ncol = 1) +
  plot_annotation(
    theme = theme(
      plot.title = element_text(size = 16, face = "bold"),
      plot.caption = element_text(size = 8)
    )
  )

print(combined)

ggsave("exp3_learn_plot.png", learn_plot_exp3, width = 6, height = 4, dpi = 300)

```


```{r t_plot_explicit, cache=F, fig.width=4, fig.height=6}

make_phase_plot_explicit <- function(df,
                                    max_trial,
                                    plot_subtitle = NULL,
                                    plot_title = NULL,
                                    bin_width = 5) {
  bins <- make_bins(max_trial = max_trial, bin_width = bin_width)
  breaks <- bins$breaks
  bin_labels <- bins$labels
  n_bins <- length(bin_labels)
  vline_x <- vline_geom(n_bins)
  
  df_long <- df %>%
    pivot_longer(
      cols = c(chose_onlystones, chose_onlystones_p2),
      names_to = "phase_raw",
      values_to = "chose_onlystones"
    ) %>%
    mutate(
      phase = case_when(
        phase_raw == "chose_onlystones"    ~ "part1",
        phase_raw == "chose_onlystones_p2" ~ "part2",
        TRUE ~ NA_character_
      ),
      phase = factor(phase, levels = c("part1", "part2")),
      bin = cut(
        trial_num,
        breaks = breaks,
        labels = bin_labels,
        include.lowest = TRUE,
        right = TRUE
      ),
      bin = factor(bin, levels = bin_labels)
    ) %>%
    filter(!is.na(phase))
  
  by_subj_bias <- df_long %>%
    group_by(subj_id, phase, bin) %>%
    summarise(bias = mean(chose_onlystones, na.rm = TRUE), .groups = "drop")
  
  group_summary_sub <- by_subj_bias %>%
    group_by(phase, bin) %>%
    summarise(
      m  = mean(bias, na.rm = TRUE),
      se = sd(bias, na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )
  
  p <- ggplot(group_summary_sub,
              aes(x = bin, y = m,
                  color = phase,
                  shape = phase,
                  group = phase)) +
    geom_hline(yintercept = 1/3, linewidth = 0.3) +
    geom_line(linewidth = 0.7, position = pd) +
    geom_point(size = 2, position = pd) +
    geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0.2, position = pd, linetype = "solid") +
   scale_color_manual(
  values = phase_pal,
  name = NULL,
  labels = c(
    "part1" = "Collect the gems!",
    "part2" = "Repeat your guesses!"
  )
) +
scale_shape_discrete(
  name = NULL,
  labels = c(
    "part1" = "Collect the gems!",
    "part2" = "Repeat your guesses!"
  )
) +
scale_linetype_discrete(
  name = NULL,
  labels = c(
    "part1" = "Collect the gems!",
    "part2" = "Repeat your guesses!"
  )
) +
    labs(x = "Trial Bin", y = "P(*stones only*)", title = plot_title, subtitle = plot_subtitle) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(size = 11, face = "bold"),
          plot.subtitle = element_text(size = 9)) +
    scale_y_continuous(limits = c(0, 0.4)) + 
     theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
    axis.title.x = element_markdown(size = 10),
    axis.title.y = element_markdown(size = 10)  ,
    legend.position = 'none'
  ) 
  
  return(p)
}


make_sig_layer <- function(ttest_df, y = 0.15, pad = 0.35, colour = "red", linewidth = 0.8) {
  sig_bins <- ttest_df %>%
    filter(!is.na(p.value), p.value < 0.05) %>%
    mutate(pos = as.integer(bin)) %>%
    arrange(pos)

  if (nrow(sig_bins) == 0) {
    return(geom_blank())
  }

  runs <- split(sig_bins, cumsum(c(1, diff(sig_bins$pos) != 1)))

  seg_df <- map_dfr(runs, function(df_run) {
    start_pos <- min(df_run$pos)
    end_pos   <- max(df_run$pos)
    tibble(
      x = start_pos - pad,
      xend = end_pos + pad,
      y = y,
      yend = y
    )
  })

  geom_segment(
    data = seg_df,
    aes(x = x, xend = xend, y = y, yend = yend),
    inherit.aes = FALSE,
    color = colour,
    linewidth = linewidth,
    lineend = "round"
  )
}


plot_exp1 <- make_phase_plot_explicit(
  two_parts_df_exp1 %>% filter(!is_catch_trial, mentioned_rule),
  max_trial = 60,
  plot_title = NULL,
  plot_subtitle = "Exp. 1"
) +
  make_sig_layer(results_exp1_explicit, y = 0)

plot_exp2 <- make_phase_plot_explicit(
  two_parts_df_exp2 %>% filter(!is_catch_trial, mentioned_rule),
  max_trial = 50,
  plot_title = NULL,
  plot_subtitle = "Exp. 2"
) +
  make_sig_layer(results_exp2_explicit, y = 0)

plot_exp3 <- make_phase_plot_explicit(
  two_parts_df_exp3 %>% filter(!is_catch_trial, mentioned_rule),
  max_trial = 50,
  plot_title = NULL,
  plot_subtitle = "Exp. 3"
) +
  make_sig_layer(results_exp3_explicit, y = 0)
combined <- plot_exp1 / plot_exp2 / plot_exp3 + 
  plot_layout(guides = "collect", ncol = 1) +
  plot_annotation(
    theme = theme(
      plot.title = element_text(size = 16, face = "bold"),
      plot.caption = element_text(size = 8)
    )
  )

print(combined)

```


```{r guess_accuracy_plot, cache=F}

exp1_accuracy <- two_parts_df_exp1 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    acc = mean(is_consistent)
  ) %>%
  mutate(chose_gems = factor(chose_gems)) %>%
  group_by(chose_gems) %>%
  summarise(
    mean_acc = mean(acc),
     se_acc   = sd(acc, na.rm = TRUE) / sqrt(n()),
    n        = n()
    
  )

exp2_accuracy <- two_parts_df_exp2 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    acc = mean(is_consistent)
  ) %>%
  mutate(chose_gems = factor(chose_gems)) %>%
  group_by(chose_gems) %>%
  summarise(
    mean_acc = mean(acc),
     se_acc   = sd(acc, na.rm = TRUE) / sqrt(n()),
    n        = n()
    
  )

exp3_accuracy <- two_parts_df_exp3 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    acc = mean(is_consistent),
   
  )


exp1_acc_wide <- two_parts_df_exp1 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    subj_acc = mean(is_consistent)
  ) %>%
  pivot_wider(names_from = chose_gems, values_from = subj_acc)%>%
  filter(!is.na(`TRUE`) & !is.na(`FALSE`))

exp2_acc_wide <- two_parts_df_exp2 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    subj_acc = mean(is_consistent)
  ) %>%
  pivot_wider(names_from = chose_gems, values_from = subj_acc)%>%
  filter(!is.na(`TRUE`) & !is.na(`FALSE`))

exp1_acc_t <- t.test(exp1_acc_wide$`TRUE`, exp1_acc_wide$`FALSE`, paired=T)
exp2_acc_t <- t.test(exp2_acc_wide$`TRUE`, exp2_acc_wide$`FALSE`, paired=T)

fmt_p <- function(p) {
  if (is.na(p)) return(NA_character_)
  if (p < 0.001) return("p < .001")
  sprintf("p = %.3f", p)
}
exp1_p <- fmt_p(exp1_acc_t$p.value)
exp2_p <- fmt_p(exp2_acc_t$p.value)

x1 <- 1
x2 <- 2
y_top <- 0.55
y <- y_top - 0.03                
cap_h <- 0.03                  
text_y <- y_top - 0.08





premembered_exp1_plot <- ggplot(exp1_accuracy, aes(x = chose_gems, y = mean_acc)) +
  geom_errorbar(
    aes(ymin = mean_acc - se_acc, ymax = mean_acc + se_acc),
    width = 0.12,
    linewidth = 0.8
  ) +
  geom_point(
    aes(color = chose_gems, fill = chose_gems),
    shape = 23,            # filled diamond
    size = 3,
  ) +
  geom_hline(yintercept = 1/3, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_y_continuous(
    limits = c(0.3, 0.8),      
    breaks = seq(0.1, 1.1, by = 0.2),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    x = "Part 1 Outcome",
    y = "P(*remembered*)"
  ) +
  scale_x_discrete(
  labels = c(
    `FALSE` = "Stones",
    `TRUE`  = "Gems"
  )
) + 
  scale_color_manual(
    values = c(
      `TRUE`  = "#22C55E",
      `FALSE` = "grey70"
    )
  ) +
  scale_fill_manual(
  values = c(
    `TRUE`  = "#22C55E",
    `FALSE` = "grey70"
  )
) + 
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
    legend.position = "none",
    axis.title.y = element_markdown(size = 14),
    axis.title.x = element_markdown(size = 14),
    axis.text = element_text(size = 12)
  ) + 
  geom_segment(aes(x = x1, xend = x2, y = y, yend = y), inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = x1, xend = x1, y = y, yend = y + cap_h), inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = x2, xend = x2, y = y, yend = y + cap_h), inherit.aes = FALSE, linewidth = 0.6) +
  annotate("text", x = (x1 + x2) / 2, y = text_y, label = exp1_p, size = 5, vjust = 0)

premembered_exp2_plot <- ggplot(exp2_accuracy, aes(x = chose_gems, y = mean_acc)) +
  geom_errorbar(
    aes(ymin = mean_acc - se_acc, ymax = mean_acc + se_acc),
    width = 0.12,
    linewidth = 0.8
  ) +
  geom_point(
    aes(color = chose_gems, fill = chose_gems),
    shape = 23,            # filled diamond
    size = 3
  ) +
  geom_hline(yintercept = 1/3, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_y_continuous(
    limits = c(0.3, 0.8),           # match your example; change if you prefer 0-1
    breaks = seq(0.1, 1.1, by = 0.2),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    x = "Part 1 Outcome",
    y = "P(*remembered*)"
  ) +
  scale_color_manual(
    values = c(
      `TRUE`  = "#22C55E",
      `FALSE` = "grey70"
    )
  ) +
  scale_fill_manual(
  values = c(
    `TRUE`  = "#22C55E",
    `FALSE` = "grey70"
  )
) + 
  scale_x_discrete(
  labels = c(
    `FALSE` = "Stones",
    `TRUE`  = "Gems"
  )
)  +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
    legend.position = "none",
    axis.title.y = element_markdown(size = 14),
    axis.title.x = element_markdown(size = 14),
    axis.text = element_text(size = 12)
  )+ 
  geom_segment(aes(x = x1, xend = x2, y = y, yend = y), inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = x1, xend = x1, y = y, yend = y + cap_h), inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = x2, xend = x2, y = y, yend = y + cap_h), inherit.aes = FALSE, linewidth = 0.6) +
  annotate("text", x = (x1 + x2) / 2, y = text_y, label = exp2_p, size = 5, vjust = 0)


print(premembered_exp1_plot)
print(premembered_exp2_plot)

```






```{r combined_plots, fig.width = 8, fig.height = 6,  cache=F}
library(patchwork) 

exp1_inconsistent <- two_parts_df_exp1 %>%
  filter(!is_consistent, !is_catch_trial, chose_gems) %>%
  group_by(subj_id) %>%
  summarise(pstone = mean(chose_onlystones_p2))

exp2_inconsistent <- two_parts_df_exp2 %>%
  filter(!is_consistent, !is_catch_trial, chose_gems) %>%
  group_by(subj_id) %>%
  summarise(pstone = mean(chose_onlystones_p2))

exp3_inconsistent <- two_parts_df_exp3 %>%
  filter(!is_consistent, !is_catch_trial, chose_gems) %>%
  group_by(subj_id) %>%
  summarise(pstone = mean(chose_onlystones_p2))

library(dplyr)
library(ggplot2)
library(ggtext) 

combined <- bind_rows(
  exp1_inconsistent %>% mutate(exp = "1"),
  exp2_inconsistent %>% mutate(exp = "2"),
  exp3_inconsistent %>% mutate(exp = "3")
) %>%
  mutate(exp = factor(exp, levels = c("1", "2", "3")))
  


summary_df <- combined %>%
  group_by(exp) %>%
  summarise(
    mean_pstone = mean(pstone, na.rm = TRUE),
    se_pstone   = sd(pstone, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )


asterisk_df_pstones <- summary_df %>%
  filter(exp != 3) %>%          
  mutate(
    label = "***",
    y = mean_pstone + se_pstone + 0.01  
  )


pstones_only_plot <-ggplot(summary_df, aes(x = exp, y = mean_pstone)) +
  geom_errorbar(
    aes(ymin = mean_pstone - se_pstone,
        ymax = mean_pstone + se_pstone),
    width = 0.12,
    linewidth = 0.8
  ) +
  geom_point(
    shape = 23,           
    size = 3,
    fill = "black",
    color = "black"
  ) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_y_continuous(
    limits = c(0.3, 0.8),
    breaks = seq(0.1, 1.1, by = 0.1),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    x = "Experiment",
    y = "P(*stones only*)"
  ) +
  theme_minimal(base_size = 12) +
   theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
  ) + 
    theme(legend.position = "none",
        axis.title.y = element_markdown(size = 14),
        axis.title.x = element_markdown(size = 14),
        axis.text = element_text(size = 12)) + 
  geom_text(
    data = asterisk_df_pstones,
    aes(x = exp, y = y, label = label),
    inherit.aes = FALSE,
    size = 6,
    vjust = 0,
    fontface = 'bold'
  )





higher_df_exp1 <- two_parts_df_exp1 %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  mutate(
    num_chosen = case_when(
      response_position_p2 == "middle" ~ middle_number,
      response_position_p2 == "left" ~ left_number,
      response_position_p2 == "right" ~ right_number
    ),
    num_unchosen = case_when(
      response_position_p2 == "middle" & response_position == "left"  ~ right_number,
      response_position_p2 == "middle" & response_position == "right" ~ left_number,
      
      response_position_p2 == "right" & response_position == "left"   ~ middle_number,
      response_position_p2 == "right" & response_position == "middle" ~ left_number,
      
      response_position_p2 == "left" & response_position == "middle" ~ right_number,
      response_position_p2 == "left" & response_position == "right"  ~ middle_number,
    ))  %>%
  filter(num_chosen != num_unchosen) %>%
  mutate(
    higher = num_chosen>num_unchosen
  ) 

higher_df_exp2 <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  mutate(
    num_chosen = case_when(
      response_position_p2 == "middle" ~ middle_number,
      response_position_p2 == "left" ~ left_number,
      response_position_p2 == "right" ~ right_number
    ),
    num_unchosen = case_when(
      response_position_p2 == "middle" & response_position == "left"  ~ right_number,
      response_position_p2 == "middle" & response_position == "right" ~ left_number,
      
      response_position_p2 == "right" & response_position == "left"   ~ middle_number,
      response_position_p2 == "right" & response_position == "middle" ~ left_number,
      
      response_position_p2 == "left" & response_position == "middle" ~ right_number,
      response_position_p2 == "left" & response_position == "right"  ~ middle_number,
    ))  %>%
  filter(num_chosen != num_unchosen) %>%
  mutate(
    higher = num_chosen>num_unchosen
  ) 

higher_df_exp3 <- two_parts_df_exp3 %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  mutate(
    num_chosen = case_when(
      response_position_p2 == "middle" ~ middle_number,
      response_position_p2 == "left" ~ left_number,
      response_position_p2 == "right" ~ right_number
    ),
    num_unchosen = case_when(
      response_position_p2 == "middle" & response_position == "left"  ~ right_number,
      response_position_p2 == "middle" & response_position == "right" ~ left_number,
      
      response_position_p2 == "right" & response_position == "left"   ~ middle_number,
      response_position_p2 == "right" & response_position == "middle" ~ left_number,
      
      response_position_p2 == "left" & response_position == "middle" ~ right_number,
      response_position_p2 == "left" & response_position == "right"  ~ middle_number,
    ))  %>%
  filter(num_chosen != num_unchosen) %>%
  mutate(
    higher = num_chosen>num_unchosen
  )  

library(dplyr)
library(ggplot2)

per_subj_higher_exp1 <- higher_df_exp1 %>%
  group_by(subj_id) %>%
  summarise(p_higher = mean(higher, na.rm = TRUE), .groups = "drop")

per_subj_higher_exp2 <- higher_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(p_higher = mean(higher, na.rm = TRUE), .groups = "drop")

per_subj_higher_exp3 <- higher_df_exp3 %>%
  group_by(subj_id) %>%
  summarise(p_higher = mean(higher, na.rm = TRUE), .groups = "drop")

combined_higher <- bind_rows(
  per_subj_higher_exp1 %>% mutate(exp = "1"),
  per_subj_higher_exp2 %>% mutate(exp = "2"),
  per_subj_higher_exp3 %>% mutate(exp = "3")
) %>%
  mutate(exp = factor(exp, levels = c("1", "2", "3")))

summary_higher <- combined_higher %>%
  group_by(exp) %>%
  summarise(
    mean_ph = mean(p_higher, na.rm = TRUE),
    se_ph   = sd(p_higher, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )

asterisk_df_phigher <- summary_higher %>%        
  mutate(
    label = "***",
    y = mean_ph + se_ph + 0.01  
  )

phigher_number_plot <-ggplot(summary_higher, aes(x = exp, y = mean_ph)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", linewidth = 0.8) +     
  geom_errorbar(aes(ymin = mean_ph - se_ph, ymax = mean_ph + se_ph),
                width = 0.12, linewidth = 0.8) +
  geom_point(shape = 23, size = 3, fill = "black", color = "black") +                    
  scale_y_continuous(limits = c(0.3, 0.8), breaks = seq(0.1, 1.1, 0.1), expand = expansion(mult = c(0, 0.08))) +
  labs(
    x = "Experiment",
    y = "P(*higher number*)",
  ) +
  theme_minimal(base_size = 12) +
   theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
  )  + 
    theme(legend.position = "none",
        axis.title.y = element_markdown(size = 14),
        axis.title.x = element_markdown(size = 14),
        axis.text = element_text(size = 12)) + 
  geom_text(
    data = asterisk_df_phigher,
    aes(x = exp, y = y, label = label),
    inherit.aes = FALSE,
    size = 6,
    vjust = 0,
    fontface = 'bold'
  )




library(lme4)
library(ggtext) 

pgems_exp1 <- two_parts_df_exp1 %>%
  filter(!is_catch_trial, !chose_gems, !is_consistent, chose_onlystones)

pgems_exp2 <- two_parts_df_exp2 %>%
  filter(!is_catch_trial, !chose_gems, !is_consistent, chose_onlystones)


library(dplyr)
library(ggplot2)


per_subj_pgems_exp1 <- pgems_exp1 %>%
  group_by(subj_id) %>%
  summarise(p_gems = mean(chose_gems_p2, na.rm = TRUE), .groups = "drop")

per_subj_pgems_exp2 <- pgems_exp2 %>%
  group_by(subj_id) %>%
  summarise(p_gems = mean(chose_gems_p2, na.rm = TRUE), .groups = "drop")

combined_pgems <- bind_rows(
  per_subj_pgems_exp1 %>% mutate(exp = "1"),
  per_subj_pgems_exp2 %>% mutate(exp = "2"),
) %>%
  mutate(exp = factor(exp, levels = c("1", "2")))


summary_pgems <- combined_pgems %>%
  group_by(exp) %>%
  summarise(
    mean_pg = mean(p_gems, na.rm = TRUE),
    se_pg   = sd(p_gems, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )

asterisk_df_pgems <- summary_pgems %>%
  mutate(
    label = case_when(
      exp == 1 ~ "***",
      exp == 2 ~ "",
      TRUE ~ NA_character_
    ),
    y = mean_pg + se_pg + 0.01
  ) %>%
  filter(!is.na(label))   


pgems_plot<- ggplot(summary_pgems, aes(x = exp, y = mean_pg)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_errorbar(aes(ymin = mean_pg - se_pg, ymax = mean_pg + se_pg),
                width = 0.12, linewidth = 0.8) +
  geom_point(shape = 23, size = 3, fill = "black", color = "black") +
  scale_y_continuous(limits = c(0.3, 0.8), breaks = seq(0.1, 1.1, 0.1), expand = expansion(mult = c(0, 0.08))) +
  labs(
    x = "Experiment",
    y = "P(*gems*)",
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
  ) + 
    theme(legend.position = "none",
        axis.title.y = element_markdown(size = 14),
        axis.title.x = element_markdown(size = 14),
        axis.text = element_text(size = 12)) + 
  geom_text(
  data = asterisk_df_pgems,
  aes(x = exp, y = y, label = label),
  inherit.aes = FALSE,
  size = 6,
  fontface = "bold",
  vjust = 0
)


library(dplyr)
library(ggplot2)
library(ggtext)

acc_subj <- two_parts_df_exp3 %>%
  filter(!is_catch_trial) %>%
  mutate(
    gem_chosen_by = case_when(
      chose_gems ~ "Self",
      computer_chose_gems ~ "Computer",
      !chose_gems & !computer_chose_gems ~ "None"
    ),
    gem_chosen_by = factor(gem_chosen_by, levels = c("Self", "Computer", "None"))
  ) %>%
  group_by(subj_id, gem_chosen_by) %>%
  summarise(mean_acc = mean(is_consistent, na.rm = TRUE), .groups = "drop")

acc_summary <- acc_subj %>%
  group_by(gem_chosen_by) %>%
  summarise(
    mean_acc_mean = mean(mean_acc, na.rm = TRUE),
    se_acc = sd(mean_acc, na.rm = TRUE) / sqrt(n()),
    n_acc = n(),
    .groups = "drop"
  )

conf_subj <- two_parts_df_exp3 %>%
  filter(!is_catch_trial, is_consistent) %>%
  mutate(
    gem_chosen_by = case_when(
      chose_gems ~ "Self",
      computer_chose_gems ~ "Computer",
      !chose_gems & !computer_chose_gems ~ "None"
    ),
    gem_chosen_by = factor(gem_chosen_by, levels = c("Self", "Computer", "None"))
  ) %>%
  group_by(subj_id, gem_chosen_by) %>%
  summarise(mean_conf = mean(confidence, na.rm = TRUE), .groups = "drop")

conf_summary <- conf_subj %>%
  group_by(gem_chosen_by) %>%
  summarise(
    mean_conf_mean = mean(mean_conf, na.rm = TRUE),
    se_conf = sd(mean_conf, na.rm = TRUE) / sqrt(n()),
    n_conf = n(),
    .groups = "drop"
  )

summary_combined <- full_join(acc_summary, conf_summary, by = "gem_chosen_by")

conf_to_left <- function(conf)  0.6 + (conf - 60) * 0.01
left_to_conf <- function(y)  60 + (y - 0.6) * 100

summary_combined <- summary_combined %>%
  mutate(
    mean_conf_left = conf_to_left(mean_conf_mean),
    se_conf_left = se_conf * 0.01  
  )


library(dplyr)
library(tidyr)
library(ggplot2)
library(ggtext)

dodge <- position_dodge(width = 0.6)

plotting_df <- bind_rows(
  summary_combined %>%
    transmute(
      gem_chosen_by,
      measure = "Reconstruction",
      mean_left = mean_acc_mean,
      se_left   = se_acc
    ),
  summary_combined %>%
    transmute(
      gem_chosen_by,
      measure = "Confidence",
      mean_left = mean_conf_left,
      se_left   = se_conf_left
    )
) %>%
  mutate(
    gem_chosen_by = factor(gem_chosen_by, levels = c("Self", "Computer", "None")),
    measure = factor(measure, levels = c("Reconstruction", "Confidence"))
  )

library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)

make_bins <- function(max_trial, bin_width = 5) {
  breaks <- seq(0.5, max_trial + 0.5, by = bin_width)
  starts <- seq(1, max_trial, by = bin_width)
  ends   <- pmin(starts + bin_width - 1, max_trial)
  labels <- paste0(starts, "–", ends)
  list(breaks = breaks, labels = labels)
}


make_phase_plot_dynamic <- function(df,
                                    max_trial,
                                    plot_subtitle = NULL,
                                    plot_title = NULL,
                                    bin_width = 5) {
  bins <- make_bins(max_trial = max_trial, bin_width = bin_width)
  breaks <- bins$breaks
  bin_labels <- bins$labels
  
  df_long <- df %>%
    pivot_longer(
      cols = c(chose_onlystones, chose_onlystones_p2),
      names_to = "phase_raw",
      values_to = "chose_onlystones"
    ) %>%
    mutate(
      phase = case_when(
        phase_raw == "chose_onlystones"    ~ "part1",
        phase_raw == "chose_onlystones_p2" ~ "part2",
        TRUE ~ NA_character_
      ),
      phase = factor(phase, levels = c("part1", "part2")),
      bin = cut(
        trial_num,
        breaks = breaks,
        labels = bin_labels,
        include.lowest = TRUE,
        right = TRUE
      ),
      bin = factor(bin, levels = bin_labels)
    ) %>%
    filter(!is.na(phase))
  
  by_subj_bias <- df_long %>%
    group_by(subj_id, phase, bin) %>%
    summarise(bias = mean(chose_onlystones, na.rm = TRUE), .groups = "drop")
  
  group_summary_sub <- by_subj_bias %>%
    group_by(phase, bin) %>%
    summarise(
      m  = mean(bias, na.rm = TRUE),
      se = sd(bias, na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )
  
  p <- ggplot(group_summary_sub,
              aes(x = bin, y = m,
                  color = phase,
                  shape = phase,
                  group = phase)) +
    geom_hline(yintercept = 1/3, linewidth = 0.3) +
    geom_line(linewidth = 0.7) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0.2) +
    labs(x = "Trial Bin", y = "P(*stones only*)", title = plot_title, subtitle = plot_subtitle) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(size = 11, face = "bold"),
          plot.subtitle = element_text(size = 9),
          legend.position = "none") +
    scale_y_continuous(limits = c(0.15, 0.4))
  
  return(p)
}


plot_left <- plotting_df %>%
  filter(measure == "Reconstruction") %>%
  ggplot(aes(x = gem_chosen_by, y = mean_left)) +
  geom_errorbar(aes(ymin = mean_left - se_left, ymax = mean_left + se_left),
                width = 0.12, linewidth = 0.8, color = "black", position = position_dodge(width = 0.2)) +
  geom_point(shape = 23, size = 3, fill = "black", color = "black", position = position_dodge(width = 0.2)) +
  scale_y_continuous(limits = c(0.6, 0.8), breaks = seq(0.6, 0.8, by = 0.05)) +
  labs(x = "Gems in Part 1 Chosen By:", y = "P(*remembered*)") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title.y = element_markdown(size = 14),
        axis.title.x = element_markdown(size = 14),
        axis.text = element_text(size = 12))


plot_right <- plotting_df %>%
  filter(measure == "Confidence") %>%
  ggplot(aes(x = gem_chosen_by, y = mean_left)) +
  geom_errorbar(aes(ymin = mean_left - se_left, ymax = mean_left + se_left),
                width = 0.12, linewidth = 0.8, color = "black", position = position_dodge(width = 0.2)) +
  geom_point(shape = 23, size = 3, fill = "black", color = "black", position = position_dodge(width = 0.2)) +
  scale_y_continuous(
    name = "Confidence",
    limits = c(0.6, 0.8),
    breaks = seq(0.6, 0.8, by = 0.05),
    labels = function(x) {
      if (exists("left_to_conf") && is.function(left_to_conf)) left_to_conf(x) else (x*150 - 30)
    }) +
  labs(x = "Gems in Part 1 Chosen By:") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title.y = element_markdown(size = 14),
        axis.title.x = element_markdown(size = 14),
        axis.text = element_text(size = 12))

acc_wide <- acc_subj %>%
  filter(gem_chosen_by %in% c("Self", "Computer", "None")) %>%
  pivot_wider(names_from = gem_chosen_by, values_from = mean_acc)

acc_ttest_selfcomputer <- t.test(acc_wide$Self, acc_wide$Computer, paired = TRUE)
acc_ttest_computernone <- t.test(acc_wide$Computer, acc_wide$None, paired = TRUE)

conf_wide <- conf_subj %>%
  filter(gem_chosen_by %in% c("Self", "Computer", "None")) %>%
  pivot_wider(names_from = gem_chosen_by, values_from = mean_conf)

conf_ttest_selfcomputer <- t.test(conf_wide$Self, conf_wide$Computer, paired = TRUE)
conf_ttest_computernone <- t.test(conf_wide$Computer, conf_wide$None, paired = TRUE)


fmt_p <- function(p) {
  if (is.na(p)) return(NA_character_)
  if (p < 0.001) return("p < .001")
  sprintf("p = %.2f", p)
}

lab_acc_selfcomp  <- fmt_p(acc_ttest_selfcomputer$p.value)
lab_acc_compnone  <- fmt_p(acc_ttest_computernone$p.value)
lab_conf_selfcomp <- fmt_p(conf_ttest_selfcomputer$p.value)
lab_conf_compnone <- fmt_p(conf_ttest_computernone$p.value)

levs <- levels(plotting_df$gem_chosen_by)
x_self     <- which(levs == "Self")
x_computer <- which(levs == "Computer")
x_none     <- which(levs == "None")


find_left_for_conf <- function(conf_val) {
  if (exists("left_to_conf") && is.function(left_to_conf)) {
    tryCatch({
      uniroot(function(l) left_to_conf(l) - conf_val, lower = 0.55, upper = 0.85)$root
    }, error = function(e) (conf_val + 30) / 150)
  } else {
    (conf_val + 30) / 150
  }
}

acc_y_selfcomp <- 0.70
acc_y_compnone <- 0.675
conf_y_selfcomp <- find_left_for_conf(70)
conf_y_compnone <- find_left_for_conf(67.5)

bar_pad   <- 0
text_pad  <- 0.006
text_size <- 5
line_w    <- 0.6

seg_df_acc <- tibble::tibble(
  x    = c(x_self + bar_pad,  x_computer + bar_pad),
  xend = c(x_computer - bar_pad, x_none - bar_pad),
  y    = c(acc_y_selfcomp, acc_y_compnone),
  label = c(lab_acc_selfcomp, lab_acc_compnone)
)

text_df_acc <- tibble::tibble(
  x = c((x_self + x_computer)/2, (x_computer + x_none)/2),
  y = seg_df_acc$y - text_pad,
  label = seg_df_acc$label
)

seg_df_conf <- tibble::tibble(
  x    = c(x_self + bar_pad,  x_computer + bar_pad),
  xend = c(x_computer - bar_pad, x_none - bar_pad),
  y    = c(conf_y_selfcomp, conf_y_compnone),
  label = c(lab_conf_selfcomp, lab_conf_compnone)
)

text_df_conf <- tibble::tibble(
  x = c((x_self + x_computer)/2, (x_computer + x_none)/2),
  y = seg_df_conf$y - text_pad,
  label = seg_df_conf$label
)

cap_half   <- 0.01
cap_height <- 0.005


premembered_exp3_plot <- plot_left +
  geom_segment(
    data = seg_df_acc,
    aes(x = x, xend = xend, y = y, yend = y),
    inherit.aes = FALSE,
    color = "black",
    linewidth = line_w,
    lineend = "butt"
  ) +
  geom_segment(
  data = seg_df_acc,
  aes(x = x, xend = x,
      y = y, yend = y + cap_height),
  inherit.aes = FALSE,
  color = "black",
  linewidth = line_w
) +
geom_segment(
  data = seg_df_acc,
  aes(x = xend, xend = xend,
      y = y, yend = y + cap_height ),
  inherit.aes = FALSE,
  color = "black",
  linewidth = line_w
)+
  geom_text(
    data = text_df_acc,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    size = text_size,
    vjust = 1
  )


conf_exp3_plot <- plot_right +
  geom_segment(
    data = seg_df_conf,
    aes(x = x, xend = xend, y = y, yend = y),
    inherit.aes = FALSE,
    color = "black",
    linewidth = line_w,
    lineend = "butt"
  ) +
  geom_segment(
  data = seg_df_acc,
  aes(x = x, xend = x,
      y = y, yend = y + cap_height),
  inherit.aes = FALSE,
  color = "black",
  linewidth = line_w
) +
geom_segment(
  data = seg_df_acc,
  aes(x = xend, xend = xend,
      y = y, yend = y + cap_height ),
  inherit.aes = FALSE,
  color = "black",
  linewidth = line_w
)+
  geom_text(
    data = text_df_conf,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    size = text_size,
    vjust = 1
  )


plot <- (pstones_only_plot + phigher_number_plot + pgems_plot + plot_layout(axis_titles = "collect")) #/ (premembered_exp3_plot + conf_exp3_plot + plot_layout(axis_titles = "collect"))

print(plot)

ggsave("biases.png", plot, width = 8, height = 3, dpi = 300)



```



```{r conf_exp2_plot, cache=F}

library(dplyr)
library(ggplot2)

conf_summary <- two_parts_df_exp2 %>%
  filter(!is_catch_trial, is_consistent) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(mean_conf_subj = mean(confidence, na.rm = TRUE)) %>%
  mutate(chose_gems = factor(chose_gems)) %>%
  group_by(chose_gems) %>%
  summarise(
    mean_conf = mean(mean_conf_subj),
    se_conf = sd(mean_conf_subj)/sqrt(n()),
    n = n()
  ) 


conf_subj <- two_parts_df_exp2 %>%
  filter(!is_catch_trial, is_consistent) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(mean_conf_subj = mean(confidence, na.rm = TRUE), .groups = "drop")


conf_wide <- conf_subj %>%
  pivot_wider(names_from = chose_gems, values_from = mean_conf_subj) %>%
  filter(!is.na(`TRUE`) & !is.na(`FALSE`))

conf_t <- t.test(conf_wide$`TRUE`, conf_wide$`FALSE`, paired = TRUE)


fmt_p <- function(p) {
  if (is.na(p)) return(NA_character_)
  if (p < 0.001) return("p < .001")
  sprintf("p = %.3f", p)
}
p_lab <- fmt_p(conf_t$p.value)

x1 <- 1
x2 <- 2
y_top <- 55
y <- y_top - 2
cap_h <- 0.8  
text_y <- y - 1.5



conf_exp2_plot <- ggplot(conf_summary, aes(x = chose_gems, y = mean_conf)) +
  geom_errorbar(aes(ymin = mean_conf - se_conf, ymax = mean_conf + se_conf),
                width = 0.12, linewidth = 0.8) +
  geom_point(aes(color = chose_gems, fill = chose_gems), shape = 23, size = 3) +
  scale_y_continuous(
    name = "Confidence",
    limits = c(50, 70),
    breaks = seq(50, 70, by = 10)
  ) +
  labs(x = "Part 1 Outcome") +
  scale_color_manual(
    values = c(
      `TRUE`  = "#22C55E",
      `FALSE` = "grey70"
    )
  ) +
  scale_fill_manual(
  values = c(
    `TRUE`  = "#22C55E",
    `FALSE` = "grey70"
  )
) + 
  scale_x_discrete(
  labels = c(
    `FALSE` = "Stones",
    `TRUE`  = "Gems"
  )
) + 
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 14),
    axis.title.x = element_markdown(size = 14),
    axis.text = element_text(size = 12)
  )

conf_exp2_plot <- conf_exp2_plot +
  geom_segment(aes(x = x1, xend = x2, y = y, yend = y), inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = x1, xend = x1, y = y, yend = y + cap_h), inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = x2, xend = x2, y = y, yend = y + cap_h), inherit.aes = FALSE, linewidth = 0.6) +
  annotate("text", x = (x1 + x2) / 2, y = text_y, label = p_lab, size = 5, vjust = 0)


conf_exp2_plot


```


```{r conf_more, cache=F,  fig.width=5, fig.height = 4}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggtext) 


conf_subj_all <- two_parts_df_exp2 %>%
  filter(!is_catch_trial) %>%
  mutate(correctness = ifelse(is_consistent, "Correct", "Incorrect")) %>%
  group_by(subj_id, correctness, chose_gems) %>%
  summarise(mean_conf_subj = mean(confidence, na.rm = TRUE), .groups = "drop")

gem_conf <- two_parts_df_exp2 %>%
  group_by(subj_id, chose_gems) %>%
  filter(!is_catch_trial) %>%
  summarise(
    mean_conf = mean(confidence)
  )

gem_conf_wide <- gem_conf %>%
  pivot_wider(
    names_from = chose_gems,
    values_from = mean_conf,
    names_prefix = "chose_gems_"
  )

gem_ttest <- t.test(
  gem_conf_wide$chose_gems_TRUE,
  gem_conf_wide$chose_gems_FALSE,
  paired = TRUE
)


conf_summary_all <- conf_subj_all %>%
  group_by(correctness, chose_gems) %>%
  summarise(
    mean_conf = mean(mean_conf_subj),
    se_conf = sd(mean_conf_subj) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )


conf_subj_correct <- conf_subj_all %>%
  filter(correctness == "Correct") %>%
  pivot_wider(names_from = chose_gems, values_from = mean_conf_subj) %>%
  filter(!is.na(`TRUE`) & !is.na(`FALSE`))
conf_t_correct <- t.test(conf_subj_correct$`TRUE`, conf_subj_correct$`FALSE`, paired = TRUE)




conf_subj_incorrect <- conf_subj_all %>%
  filter(correctness == "Incorrect") %>%
  pivot_wider(names_from = chose_gems, values_from = mean_conf_subj) %>%
  filter(!is.na(`TRUE`) & !is.na(`FALSE`))
conf_t_incorrect <- t.test(conf_subj_incorrect$`TRUE`, conf_subj_incorrect$`FALSE`, paired = TRUE)


fmt_p <- function(p) {
  if (is.na(p)) return(NA_character_)
  if (p < 0.001) return("p < .001")
  sprintf("p = %.3f", p)
}
p_lab_correct   <- fmt_p(conf_t_correct$p.value)
p_lab_incorrect <- fmt_p(conf_t_incorrect$p.value)

conf_summary_all <- conf_summary_all %>%
  mutate(x_level = interaction(correctness, chose_gems, sep = "_")) %>%
  mutate(x_level = factor(
    x_level,
    levels = c("Correct_FALSE", "Correct_TRUE", "Incorrect_FALSE", "Incorrect_TRUE")
  ))

x1 <- 1
x2 <- 2
x_mid <- (x1 + x2) / 2

y_correct <- 53
cap_h_correct <- 1.5
text_y_correct <- y_correct - 2.5   # below bracket

y_incorrect <- 55
cap_h_incorrect <- 1.5
text_y_incorrect <- y_incorrect + 2.5  # above bracket


bracket_correct_h  <- data.frame(correctness = "Correct", x = x1, xend = x2, y = y_correct, yend = y_correct)
bracket_correct_cap1 <- data.frame(correctness = "Correct", x = x1, xend = x1, y = y_correct, yend = y_correct + cap_h_correct)
bracket_correct_cap2 <- data.frame(correctness = "Correct", x = x2, xend = x2, y = y_correct, yend = y_correct + cap_h_correct)

bracket_incorrect_h  <- data.frame(correctness = "Incorrect", x = x1, xend = x2, y = y_incorrect, yend = y_incorrect)
bracket_incorrect_cap1 <- data.frame(correctness = "Incorrect", x = x1, xend = x1, y = y_incorrect, yend = y_incorrect - cap_h_incorrect)
bracket_incorrect_cap2 <- data.frame(correctness = "Incorrect", x = x2, xend = x2, y = y_incorrect, yend = y_incorrect - cap_h_incorrect)


conf_combined_plot <- ggplot(
  conf_summary_all,
  aes(x = factor(chose_gems), y = mean_conf)
) +
  geom_errorbar(
    aes(ymin = mean_conf - se_conf, ymax = mean_conf + se_conf),
    width = 0.12, linewidth = 0.8
  ) +
  geom_point(
    aes(color = factor(chose_gems), fill = factor(chose_gems)),
    shape = 23, size = 3
  ) +
  scale_y_continuous(
    name = "Confidence",
    limits = c(40, 70),
    breaks = seq(40, 70, by = 10)
  ) +
  labs(x = "Part 1 Outcome") +
  scale_color_manual(
    values = c(`TRUE` = "#22C55E", `FALSE` = "grey70")
  ) +
  scale_fill_manual(
    values = c(`TRUE` = "#22C55E", `FALSE` = "grey70")
  ) +
  scale_x_discrete(
    labels = c(`FALSE` = "Stones", `TRUE` = "Gems")
  ) +
  facet_wrap(
    ~ correctness,
    nrow = 1,
    labeller = labeller(
      correctness = c(
        "Correct" = "Correct",
        "Incorrect" = "Incorrect"
      )
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 14),
    axis.title.x = element_markdown(size = 14),
    axis.text = element_text(size = 12),

    strip.text = element_text(size = 14, face = "bold"),
    strip.background = element_blank()
  )


conf_combined_plot <- conf_combined_plot +
  geom_segment(
    data = data.frame(correctness = "Correct"),
    aes(x = x1, xend = x2, y = y_correct, yend = y_correct),
    inherit.aes = FALSE, linewidth = 0.6
  ) +
  geom_segment(
    data = data.frame(correctness = "Correct"),
    aes(x = x1, xend = x1, y = y_correct, yend = y_correct + cap_h_correct),
    inherit.aes = FALSE, linewidth = 0.6
  ) +
  geom_segment(
    data = data.frame(correctness = "Correct"),
    aes(x = x2, xend = x2, y = y_correct, yend = y_correct + cap_h_correct),
    inherit.aes = FALSE, linewidth = 0.6
  ) +
  geom_text(
    data = data.frame(correctness = "Correct"),
    aes(x = x_mid, y = text_y_correct, label = "***", fontface = 'bold' ),
    inherit.aes = FALSE, size = 5
  ) +
  geom_segment(
    data = data.frame(correctness = "Incorrect"),
    aes(x = x1, xend = x2, y = y_incorrect, yend = y_incorrect),
    inherit.aes = FALSE, linewidth = 0.6
  ) +
  geom_segment(
    data = data.frame(correctness = "Incorrect"),
    aes(x = x1, xend = x1, y = y_incorrect, yend = y_incorrect - cap_h_incorrect),
    inherit.aes = FALSE, linewidth = 0.6
  ) +
  geom_segment(
    data = data.frame(correctness = "Incorrect"),
    aes(x = x2, xend = x2, y = y_incorrect, yend = y_incorrect - cap_h_incorrect),
    inherit.aes = FALSE, linewidth = 0.6
  ) +
  geom_text(
    data = data.frame(correctness = "Incorrect"),
    aes(x = x_mid, y = text_y_incorrect,  label = "***", fontface = 'bold' ),
    inherit.aes = FALSE, size = 5
  )


conf_combined_plot


ggsave("conf_combined_exp2.png",
       conf_combined_plot,
       width = 5,
       height = 4,
       dpi = 300)


conf_correct <- two_parts_df_exp2 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, is_consistent) %>%
  summarise(
    mean_conf = mean(confidence, na.rm = TRUE),
    .groups = "drop"
  )

conf_wide <- conf_correct %>%
  pivot_wider(
    names_from = is_consistent,
    values_from = mean_conf,
    names_prefix = "consistent_"
  )

t.test(
  conf_wide$consistent_TRUE,
  conf_wide$consistent_FALSE,
  paired = TRUE
)

library(lme4)
library(lmerTest)  

conf_lmm <- lmer(
  confidence ~ chose_gems * is_consistent + (1 | subj_id),
  data = two_parts_df_exp2 %>% filter(!is_catch_trial)
)

anova(conf_lmm, type = 3)


```

difference between chose gems and chose stones in part 1 in part 2 correct trials `r apa_print(conf_t_correct)$statistic`

difference between chose gems and chose stones in part 1 in part 2 incorrect trials `r apa_print(conf_t_incorrect)$statistic`

difference between correct and incorrect part 2 confidence `r apa_print(t.test(conf_wide$consistent_TRUE,conf_wide$consistent_FALSE,paired = TRUE))$statistic`

t-test between correct and incorrect (irrespective of part 1 choice): `r apa_print(t.test(conf_wide$consistent_TRUE,conf_wide$consistent_FALSE,paired = TRUE))$statistic`

t-test between gems/stones in part 1, for incorrect decisions: `r apa_print(conf_t_incorrect)$statistic`

t-test between gems/stones in part 1 (irrespective of accuracy): `r apa_print(gem_ttest)$statistic`

A linear mixed-effects model was fit to predict confidence with chose_gems (Gems vs. Stones), is_consistent (Correct vs. Incorrect), and their interaction as fixed effects, and a random intercept for participant (subj_id). The ANOVA on the mixed model tested the interaction between Part 1 decision and Part 2 accuracy. The interaction of Part 1 decision and Part 2 accuracy was 'reported below', `apa_print(anova(conf_lmm, type = 3))$statistic`) (see ANOVA table).


```{r correlation_exp1_2, fig.width=6, fig.height = 3, cache=F}


library(ggplot2)
library(patchwork)
library(ggtext)

cor_label <- function(x, y) {
  ct <- cor.test(x, y, method = "pearson")
  
  p_text <- if (ct$p.value < 0.001) {
    "p < 0.001"
  } else {
    paste0("p = ", signif(ct$p.value, 3))
  }
  
  paste0(
    "r = ", round(ct$estimate, 2), "\n",
    p_text
  )
}

make_cor_plot <- function(df, x_col, y_col, title,
                          color = "black", x_lab = NULL, y_lab = NULL) {
  
  label <- cor_label(x, y)  
  
  ggplot(df, aes(x = .data[[x_col]], y = .data[[y_col]])) +
    geom_point(alpha = 0.1) +
    geom_smooth(method = "lm", se = FALSE, color = color) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", alpha = 0.4) +
    annotate(
      "text",
      x = -Inf, y = Inf,
      label = label,
      hjust = -0.1, vjust = 1.1,
      size = 5
    ) +
    labs(
      title = title,
      x = ifelse(is.null(x_lab), x_col, x_lab),
      y = ifelse(is.null(y_lab), y_col, y_lab)
    ) +
    theme_minimal() +
    theme(plot.title = element_text(size = 10, hjust = 0.5),
          axis.text = element_text(size = 14)) + 
    coord_cartesian(xlim = c(0, 50), ylim = c(0, 50))
}


total_chose_gems_part1 <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(total_gems = sum(chose_gems, na.rm = TRUE))

estimate_chose_gems_part1 <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part1, na.rm = TRUE))

correlation_data_part1 <- inner_join(total_chose_gems_part1,
                                     estimate_chose_gems_part1,
                                     by = "subj_id")
label_part1 <- cor_label(
  correlation_data_part1$total_gems,
  correlation_data_part1$mean_estimate
)



total_chose_gems_part2 <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(total_chose_gems_p2 = sum(chose_gems_p2, na.rm = TRUE))

estimate_chose_gems_part2 <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part2, na.rm = TRUE))

cor_data_part2 <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(
    consistent = sum(is_consistent),
    mean_estimate_part2 = mean(number_correct_part2)
  )

total_consistent <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(total_consistent = sum(is_consistent))
estimate_part2 <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part2))

stones_part2 <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  mutate(
    stones_chosen = chose_other_p2|chose_onlystones_p2
  ) %>%
  summarise(stones= sum(stones_chosen))


correlation_data_part2 <- inner_join(total_chose_gems_part1,
                                     total_chose_gems_part2,
                                     by = "subj_id")

correlation_data_part3 <- inner_join(estimate_part2,
                                     total_chose_gems_part2,
                                     by = "subj_id")



# label_part2 <- cor_label(
#   correlation_data_part2$consistent,
#   correlation_data_part2$mean_estimate
# )

p1<- make_cor_plot(
  correlation_data_part1,
  x_col = "total_gems",
  y_col = "mean_estimate",
  title = "",
  color = "black",
  x_lab = "",
  y_lab = ""
)

p2 <- make_cor_plot(
  correlation_data_part2,
  x_col = "total_gems",
  y_col = "total_chose_gems_p2",
  title = "",
  color = "black",
  x_lab = "",
  y_lab = ""
)

library(patchwork)
library(cowplot)

exp2_cor_plot <- ggdraw(p1 + p2) +
  draw_label(
    "True Number",
    x = 0.5,
    y = 0.02,
    vjust = 0,
    size = 16,
    fontface = "bold"
  ) +
  draw_label(
    "Estimate",
    x = 0.02,
    y = 0.5,
    angle = 90,
    vjust = 1,
    size = 16,
    fontface = "bold"
  ) + 
  theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
    axis.text = element_text(size = 14)) 

ggsave("exp2_cor_plot.png", exp2_cor_plot, 
       width = 6, height = 3,
       dpi = 300)
  




  
```






```{r correlation_matrice, fig.width = 6, fig.height = 6}

library(ggtext)


cor_label <- function(x, y) {
  ct <- cor.test(x, y, method = "pearson")
  
  p_text <- if (ct$p.value < 0.001) {
    "p < 0.001"
  } else {
    paste0("p = ", signif(ct$p.value, 3))
  }
  
  paste0(
    "r = ", round(ct$estimate, 2), "\n",
    p_text
  )
}

make_cor_plot <- function(df, x_col, y_col, title,
                          color = "black", x_lab = NULL, y_lab = NULL) {

  label <- cor_label(x, y)
  
  ggplot(df, aes(x = .data[[x_col]], y = .data[[y_col]])) +
    geom_point(alpha = 0.3) +
    geom_smooth(method = "lm", se = FALSE, color = color) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", alpha = 0.4) +
    annotate(
      "text",
      x = -Inf, y = Inf,
      label = label,
      hjust = -0.1, vjust = 1.1,
      size = 5
    ) +
    labs(
      title = title,
      x = ifelse(is.null(x_lab), x_col, x_lab),
      y = ifelse(is.null(y_lab), y_col, y_lab)
    ) +
    theme_minimal() +
    theme(plot.title = element_text(size = 10, hjust = 0.5),
          axis.text = element_text(size = 14)) + 
    coord_cartesian(xlim = c(0, 50), ylim = c(0, 50))
}



estimate_chose_gems_part1 <- two_parts_df_exp3 %>%
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part1))

total_chose_gems_part1 <- two_parts_df_exp3 %>%
  group_by(subj_id) %>%
  summarise(total_chose_gems = sum(chose_gems))


estimate_chose_gems_part1 <- two_parts_df_exp3 %>%
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part1))

total_rewarded <- two_parts_df_exp3 %>%
  group_by(subj_id) %>%
  mutate(rewarded = case_when(
    chose_gems|computer_chose_gems ~ TRUE,
    TRUE~FALSE
  )) %>%
  summarise(total_rewarded = sum(rewarded))



total_consistent <- two_parts_df_exp3 %>%
  group_by(subj_id) %>%
  summarise(total_consistent = sum(is_consistent))
estimate_part2 <- two_parts_df_exp3 %>%
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part2))



total_computer_chose_gems <- two_parts_df_exp3 %>%
  group_by(subj_id) %>%
  summarise(total_computer_chose_gems = sum(computer_chose_gems))

estimate_computer_chose_gems <- two_parts_df_exp3 %>%
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part1_computer))


df1 <- inner_join(
  estimate_chose_gems_part1,
  total_chose_gems_part1,
  by = "subj_id"
)

df2 <- inner_join(
  estimate_chose_gems_part1,
  total_rewarded,
  by = "subj_id"
)

df3 <- inner_join(
  estimate_computer_chose_gems,
  total_computer_chose_gems,
  by = "subj_id"
)

df4 <- inner_join(
  estimate_computer_chose_gems,
  total_rewarded,
  by = "subj_id"
)

p1<- make_cor_plot(
  df1,
  x_col = "total_chose_gems",
  y_col = "mean_estimate",
  title = "",
  color = p1_colour,
  x_lab = "",
  y_lab = ""
)

p2 <- make_cor_plot(
  df2,
  x_col = "total_rewarded",
  y_col = "mean_estimate",
  title = "",
  color = p1_colour,
  x_lab = "",
  y_lab = ""
)

p3 <- make_cor_plot(
  df3,
  x_col = "total_computer_chose_gems",
  y_col = "mean_estimate",
  title = "",
  color = p1_colour,
  x_lab = "",
  y_lab = ""
)

p4 <- make_cor_plot(
  df4,
  x_col = "total_rewarded",
  y_col = "mean_estimate",
  title = "",
  color = p1_colour,
  x_lab = "",
  y_lab = ""
)


library(cowplot)

plot_core <- (p1 + p2) / (p3 + p4) 



final_plot <- ggdraw(plot_core) +
  draw_label(
    "True Number",
    x = 0.5,
    y = 0.02,
    vjust = 0,
    size = 16,
    fontface = "bold"
  ) +
  draw_label(
    "Estimate",
    x = 0.02,
    y = 0.5,
    angle = 90,
    vjust = 1,
    size = 16,
    fontface = "bold"
  )

print(final_plot  & theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
    axis.text = element_text(size = 14)
  ))

ggsave('cor_matrice.png', 
        final_plot  & theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
    axis.text = element_text(size = 14)
  ),
  width = 6,
  height = 6,
  dpi = 300
  )

```


## Final Plots

```{r fig1_1, cache=F, fig.width=5, fig.height=4}

library(patchwork)

fig1_1 <- learn_plot_exp1 

print(fig1_1)

ggsave("fig1_1.png", plot = fig1_1, width = 5, 
  height = 4, 
  units = "in", 
  dpi = 300)
```

```{r fig1_2, cache=F, fig.width=2.5, fig.height = 4}


exp1_accuracy <- two_parts_df_exp1 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    acc = mean(is_consistent)
  ) %>%
  mutate(chose_gems = factor(chose_gems)) %>%
  group_by(chose_gems) %>%
  summarise(
    mean_acc = mean(acc),
     se_acc   = sd(acc, na.rm = TRUE) / sqrt(n()),
    n        = n()
    
  )

exp2_accuracy <- two_parts_df_exp2 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    acc = mean(is_consistent)
  ) %>%
  mutate(chose_gems = factor(chose_gems)) %>%
  group_by(chose_gems) %>%
  summarise(
    mean_acc = mean(acc),
     se_acc   = sd(acc, na.rm = TRUE) / sqrt(n()),
    n        = n()
    
  )

exp3_accuracy <- two_parts_df_exp3 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    acc = mean(is_consistent),
   
  )


exp1_acc_wide <- two_parts_df_exp1 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    subj_acc = mean(is_consistent)
  ) %>%
  pivot_wider(names_from = chose_gems, values_from = subj_acc)%>%
  filter(!is.na(`TRUE`) & !is.na(`FALSE`))

exp2_acc_wide <- two_parts_df_exp2 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    subj_acc = mean(is_consistent)
  ) %>%
  pivot_wider(names_from = chose_gems, values_from = subj_acc)%>%
  filter(!is.na(`TRUE`) & !is.na(`FALSE`))

exp1_acc_t <- t.test(exp1_acc_wide$`TRUE`, exp1_acc_wide$`FALSE`, paired=T)
exp2_acc_t <- t.test(exp2_acc_wide$`TRUE`, exp2_acc_wide$`FALSE`, paired=T)



fmt_p <- function(p) {
  if (is.na(p)) return(NA_character_)
  if (p < 0.001) return("p < .001")
  sprintf("p = %.3f", p)
}
exp1_p <- fmt_p(exp1_acc_t$p.value)
exp2_p <- fmt_p(exp2_acc_t$p.value)

x1 <- 1
x2 <- 2
y_top <- 0.55
y <- y_top - 0.03
cap_h <- 0.03    
text_y <- y_top - 0.08





premembered_exp1_plot <- ggplot(exp1_accuracy, aes(x = chose_gems, y = mean_acc)) +
  geom_errorbar(
    aes(ymin = mean_acc - se_acc, ymax = mean_acc + se_acc),
    width = 0.12,
    linewidth = 0.8
  ) +
  geom_point(
    aes(color = chose_gems, fill = chose_gems),
    shape = 23,  
    size = 3,
  ) +
  geom_hline(yintercept = 1/3, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_y_continuous(
    limits = c(0.3, 0.8),           
    breaks = seq(0.1, 1.1, by = 0.2),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    x = "Part 1 Outcome",
    y = "P(*remembered*)"
  ) +
  scale_x_discrete(
  labels = c(
    `FALSE` = "Stones",
    `TRUE`  = "Gems"
  )
) + 
  scale_color_manual(
    values = c(
      `TRUE`  = "#22C55E",
      `FALSE` = "grey70"
    )
  ) +
  scale_fill_manual(
  values = c(
    `TRUE`  = "#22C55E",
    `FALSE` = "grey70"
  )
) + 
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
    legend.position = "none",
    axis.title.y = element_markdown(size = 14),
    axis.title.x = element_markdown(size = 14),
    axis.text = element_text(size = 12)
  ) + 
  geom_segment(aes(x = x1, xend = x2, y = y, yend = y), inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = x1, xend = x1, y = y, yend = y + cap_h), inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = x2, xend = x2, y = y, yend = y + cap_h), inherit.aes = FALSE, linewidth = 0.6) +
  annotate("text", x = (x1 + x2) / 2, y = text_y, label = "***", size = 6, vjust = 0, fontface = "bold")


fig1_2 <- premembered_exp1_plot

print(fig1_2)

ggsave("fig1_2.png", plot = fig1_2, width = 2.5, 
  height = 4, 
  units = "in", 
  dpi = 300)
```

```{r fig2_1, cache=F, fig.width=7, fig.height=4}
fig2_1 <- learn_plot_exp2

print(fig2_1)

ggsave("fig2_1.png", plot = fig2_1, width = 5, 
  height = 4, 
  units = "in", 
  dpi = 300)
```

```{r fig2_2, cache=F, fig.width=5, fig.height = 4}

library(patchwork)

exp1_accuracy <- two_parts_df_exp1 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    acc = mean(is_consistent)
  ) %>%
  mutate(chose_gems = factor(chose_gems)) %>%
  group_by(chose_gems) %>%
  summarise(
    mean_acc = mean(acc),
     se_acc   = sd(acc, na.rm = TRUE) / sqrt(n()),
    n        = n()
    
  )

exp2_accuracy <- two_parts_df_exp2 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    acc = mean(is_consistent)
  ) %>%
  mutate(chose_gems = factor(chose_gems)) %>%
  group_by(chose_gems) %>%
  summarise(
    mean_acc = mean(acc),
     se_acc   = sd(acc, na.rm = TRUE) / sqrt(n()),
    n        = n()
    
  )

exp3_accuracy <- two_parts_df_exp3 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    acc = mean(is_consistent),
   
  )


exp1_acc_wide <- two_parts_df_exp1 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    subj_acc = mean(is_consistent)
  ) %>%
  pivot_wider(names_from = chose_gems, values_from = subj_acc)%>%
  filter(!is.na(`TRUE`) & !is.na(`FALSE`))

exp2_acc_wide <- two_parts_df_exp2 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(
    subj_acc = mean(is_consistent)
  ) %>%
  pivot_wider(names_from = chose_gems, values_from = subj_acc)%>%
  filter(!is.na(`TRUE`) & !is.na(`FALSE`))

exp1_acc_t <- t.test(exp1_acc_wide$`TRUE`, exp1_acc_wide$`FALSE`, paired=T)
exp2_acc_t <- t.test(exp2_acc_wide$`TRUE`, exp2_acc_wide$`FALSE`, paired=T)


fmt_p <- function(p) {
  if (is.na(p)) return(NA_character_)
  if (p < 0.001) return("p < .001")
  sprintf("p = %.3f", p)
}
exp1_p <- fmt_p(exp1_acc_t$p.value)
exp2_p <- fmt_p(exp2_acc_t$p.value)

x1 <- 1
x2 <- 2
y_top <- 0.55
y <- y_top - 0.03                
cap_h <- 0.03                  
text_y <- y_top - 0.08

acc_y_line <- 0.55    
acc_cap_h  <- 0.02    
acc_text_y <- 0.5

premembered_exp2_plot <- ggplot(exp2_accuracy, aes(x = chose_gems, y = mean_acc)) +
  geom_errorbar(
    aes(ymin = mean_acc - se_acc, ymax = mean_acc + se_acc),
    width = 0.12,
    linewidth = 0.8
  ) +
  geom_point(
    aes(color = chose_gems, fill = chose_gems),
    shape = 23,       
    size = 3
  ) +
  geom_hline(yintercept = 1/3, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_y_continuous(
    limits = c(0.3, 0.8),
    breaks = seq(0.1, 1.1, by = 0.2),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    x = "Part 1 Outcome",
    y = "P(*remembered*)"
  ) +
  scale_color_manual(
    values = c(
      `TRUE`  = "#22C55E",
      `FALSE` = "grey70"
    )
  ) +
  scale_fill_manual(
  values = c(
    `TRUE`  = "#22C55E",
    `FALSE` = "grey70"
  )
) + 
  scale_x_discrete(
  labels = c(
    `FALSE` = "Stones",
    `TRUE`  = "Gems"
  )
)  +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_markdown(size = 12, hjust = 0.5),
    legend.position = "none",
    axis.title.y = element_markdown(size = 14),
    axis.title.x = element_markdown(size = 14),
    axis.text = element_text(size = 12)
  )+ 
  geom_segment(aes(x = 1, xend = 2, y = acc_y_line, yend = acc_y_line), 
               inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = 1, xend = 1, y = acc_y_line, yend = acc_y_line + acc_cap_h), 
               inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = 2, xend = 2, y = acc_y_line, yend = acc_y_line + acc_cap_h), 
               inherit.aes = FALSE, linewidth = 0.6) +
  annotate("text", x = 1.5, y = acc_text_y, label = "**", fontface = "bold", size = 5, vjust = 0)

library(dplyr)
library(ggplot2)

conf_summary <- two_parts_df_exp2 %>%
  filter(!is_catch_trial, is_consistent) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(mean_conf_subj = mean(confidence, na.rm = TRUE)) %>%
  mutate(chose_gems = factor(chose_gems)) %>%
  group_by(chose_gems) %>%
  summarise(
    mean_conf = mean(mean_conf_subj),
    se_conf = sd(mean_conf_subj)/sqrt(n()),
    n = n()
  ) 

conf_subj <- two_parts_df_exp2 %>%
  filter(!is_catch_trial, is_consistent) %>%
  group_by(subj_id, chose_gems) %>%
  summarise(mean_conf_subj = mean(confidence, na.rm = TRUE), .groups = "drop")


conf_wide <- conf_subj %>%
  pivot_wider(names_from = chose_gems, values_from = mean_conf_subj) %>%
  filter(!is.na(`TRUE`) & !is.na(`FALSE`))


conf_t <- t.test(conf_wide$`TRUE`, conf_wide$`FALSE`, paired = TRUE)

fmt_p <- function(p) {
  if (is.na(p)) return(NA_character_)
  if (p < 0.001) return("p < .001")
  sprintf("p = %.3f", p)
}
p_lab <- fmt_p(conf_t$p.value)

x1 <- 1
x2 <- 2
y_top <- 55
y <- y_top - 2                
cap_h <- 0.8                  
text_y <- y - 1.5


conf_exp2_plot <- ggplot(conf_summary, aes(x = chose_gems, y = mean_conf)) +
  geom_errorbar(aes(ymin = mean_conf - se_conf, ymax = mean_conf + se_conf),
                width = 0.12, linewidth = 0.8) +
  geom_point(aes(color = chose_gems, fill = chose_gems), shape = 23, size = 3) +
  scale_y_continuous(
    name = "Confidence",
    limits = c(50, 70),
    breaks = seq(50, 70, by = 10)
  ) +
  labs(x = "Part 1 Outcome") +
  scale_color_manual(
    values = c(
      `TRUE`  = "#22C55E",
      `FALSE` = "grey70"
    )
  ) +
  scale_fill_manual(
  values = c(
    `TRUE`  = "#22C55E",
    `FALSE` = "grey70"
  )
) + 
  scale_x_discrete(
  labels = c(
    `FALSE` = "Stones",
    `TRUE`  = "Gems"
  )
) + 
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 14),
    axis.title.x = element_markdown(size = 14),
    axis.text = element_text(size = 12)
  )

conf_exp2_plot <- conf_exp2_plot +
  geom_segment(aes(x = x1, xend = x2, y = y, yend = y), inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = x1, xend = x1, y = y, yend = y + cap_h), inherit.aes = FALSE, linewidth = 0.6) +
  geom_segment(aes(x = x2, xend = x2, y = y, yend = y + cap_h), inherit.aes = FALSE, linewidth = 0.6) +
  annotate("text", x = (x1 + x2) / 2, y = text_y, label = "***", fontface = "bold", size = 5, vjust = 0)

fig2_2 <- premembered_exp2_plot + conf_combined_plot +
  plot_layout(
    widths = c(1, 2),
    axis_titles = "collect"
  )


print(fig2_2)

ggsave("fig2_2.png", plot = fig2_2, width = 6, 
  height = 3.5, 
  units = "in", 
  dpi = 300)

```

t-test between gem and stone: `r apa_print(conf_t)$statistic`

```{r fig3, cache=F, fig.width = 8, fig.height = 3}
library(patchwork)
fig3 <- premembered_exp3_plot + conf_exp3_plot + plot_layout(axis_titles = "collect")

print(fig3)
         

ggsave("fig3.png", plot = fig3, width = 8, 
  height = 3, 
  units = "in", 
  dpi = 300)

```

```{r fig4, cache=F, fig.width = 8, fig.height=3}

fig4 <- pstones_only_plot + phigher_number_plot + plot_layout(axis_titles = "collect")

print(fig4)

ggsave(
  "fig4.png", 
  plot = fig4, 
  width = 8*2/3, 
  height = 3, 
  units = "in", 
  dpi = 300
)

pgems_exp1 <- two_parts_df_exp1 %>%
  filter(!is_catch_trial, !chose_gems, !is_consistent, chose_onlystones)

pgems_exp2 <- two_parts_df_exp2 %>%
  filter(!is_catch_trial, !chose_gems, !is_consistent, chose_onlystones)

per_subj_pgems_exp1 <- pgems_exp1 %>%
  group_by(subj_id) %>%
  summarise(p_gems = mean(chose_gems_p2, na.rm = TRUE), .groups = "drop")

per_subj_pgems_exp2 <- pgems_exp2 %>%
  group_by(subj_id) %>%
  summarise(p_gems = mean(chose_gems_p2, na.rm = TRUE), .groups = "drop")

model_exp1 <- glmer(
  p_gems ~ 1 + (1 | subj_id),
  data = per_subj_pgems_exp1,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa"))

model_exp2  <- glmer(
  p_gems ~ 1 + (1 | subj_id),
  data = per_subj_pgems_exp2,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa"))


```


P(gems) in trials where participant didnt chose only stones in part 1 and reconstructed incorrectly. Exp. 1 p(gems) = `r printnum(mean(per_subj_pgems_exp1$p_gems))` (`r apa_print(t.test(per_subj_pgems_exp1$p_gems, mu=0.5) )$statistic`). Exp. 2 p(gems) = `r printnum(mean(per_subj_pgems_exp2$p_gems))` (`r apa_print(t.test(per_subj_pgems_exp2$p_gems, mu=0.5) )$statistic`). 

P(gems) in trials where participant didnt chose only stones in part 1 and reconstructed incorrectly. estimated exp. 1 p(gems)  = `r plogis(round(fixef(model_exp1)[1], 2))` (β = `r printnum(round(fixef(model_exp1)[1], 2))`, 
SE = `r printnum(round(summary(model_exp1)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(model_exp1)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(model_exp1)$coefficients[1, "Pr(>|z|)"])`)

In exp. 2 p(gems) = = `r plogis(round(fixef(model_exp2)[1], 2))` (β = `r printnum(round(fixef(model_exp2)[1], 2))`, 
SE = `r printnum(round(summary(model_exp2)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(model_exp2)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(model_exp2)$coefficients[1, "Pr(>|z|)"])`)


```{r higher_analysis, cache = F}

m <- two_parts_df_exp1 %>%
  mutate(left_number = 5, middle_number = 4, right_number = 3, stoneposition = "left", response_position_p2 = "middle", response_position = "left")


check <- m %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  mutate(
    higher = case_when(
      response_position_p2 == "middle" & response_position == "left"  ~ middle_number > right_number,
      response_position_p2 == "middle" & response_position == "right" ~ middle_number > left_number,
      
      response_position_p2 == "right" & response_position == "left"   ~ right_number > middle_number,
      response_position_p2 == "right" & response_position == "middle" ~ right_number > left_number,
      
      response_position_p2 == "left" & response_position == "middle" ~ left_number > right_number,
      response_position_p2 == "left" & response_position == "right"  ~ left_number > middle_number,
      
      TRUE ~ NA
    ))

new <- m %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  mutate(
    
  )

higher_df_exp1 <- two_parts_df_exp1 %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  mutate(
    num_chosen = case_when(
      response_position_p2 == "middle" ~ middle_number,
      response_position_p2 == "left" ~ left_number,
      response_position_p2 == "right" ~ right_number
    ),
    num_unchosen = case_when(
      response_position_p2 == "middle" & response_position == "left"  ~ right_number,
      response_position_p2 == "middle" & response_position == "right" ~ left_number,
      
      response_position_p2 == "right" & response_position == "left"   ~ middle_number,
      response_position_p2 == "right" & response_position == "middle" ~ left_number,
      
      response_position_p2 == "left" & response_position == "middle" ~ right_number,
      response_position_p2 == "left" & response_position == "right"  ~ middle_number,
    ))  %>%
  filter(num_chosen != num_unchosen) %>%
  mutate(
    higher = num_chosen>num_unchosen
  )

higher_df_exp2 <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  mutate(
    num_chosen = case_when(
      response_position_p2 == "middle" ~ middle_number,
      response_position_p2 == "left" ~ left_number,
      response_position_p2 == "right" ~ right_number
    ),
    num_unchosen = case_when(
      response_position_p2 == "middle" & response_position == "left"  ~ right_number,
      response_position_p2 == "middle" & response_position == "right" ~ left_number,
      
      response_position_p2 == "right" & response_position == "left"   ~ middle_number,
      response_position_p2 == "right" & response_position == "middle" ~ left_number,
      
      response_position_p2 == "left" & response_position == "middle" ~ right_number,
      response_position_p2 == "left" & response_position == "right"  ~ middle_number,
    ))  %>%
  filter(num_chosen != num_unchosen) %>%
  mutate(
    higher = num_chosen>num_unchosen
  ) 


higher_df_exp3 <- two_parts_df_exp3 %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  mutate(
    num_chosen = case_when(
      response_position_p2 == "middle" ~ middle_number,
      response_position_p2 == "left" ~ left_number,
      response_position_p2 == "right" ~ right_number
    ),
    num_unchosen = case_when(
      response_position_p2 == "middle" & response_position == "left"  ~ right_number,
      response_position_p2 == "middle" & response_position == "right" ~ left_number,
      
      response_position_p2 == "right" & response_position == "left"   ~ middle_number,
      response_position_p2 == "right" & response_position == "middle" ~ left_number,
      
      response_position_p2 == "left" & response_position == "middle" ~ right_number,
      response_position_p2 == "left" & response_position == "right"  ~ middle_number,
    ))  %>%
  filter(num_chosen != num_unchosen) %>%
  mutate(
    higher = num_chosen>num_unchosen
  ) 




model_exp1 <- glmer(
  higher ~ 1 + (1 | subj_id),
  data = higher_df_exp1,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa"))

model_exp2 <- glmer(
  higher ~ 1 + (1 | subj_id),
  data = higher_df_exp2,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa"))


model_exp3 <- glmer(
  higher ~ 1 + (1 | subj_id),
  data = higher_df_exp2,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa"))

p_higher_exp1 <- higher_df_exp1 %>% group_by(subj_id) %>% summarise(mean_higher = mean(higher)) %>% pull(mean_higher) %>% mean()

p_higher_exp2 <- higher_df_exp2 %>% group_by(subj_id) %>% summarise(mean_higher = mean(higher)) %>% pull(mean_higher) %>% mean(na.rm=T)

p_mean_exp1 <- two_parts_df_exp1 %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  mutate(
    num_chosen = case_when(
      response_position == "middle" ~ middle_number,
      response_position == "left" ~ left_number,
      response_position == "right" ~ right_number
    )
  ) %>%
  summarise(mean_num_chosen = mean(num_chosen)) %>%
  mutate(diff = mean_num_chosen - 3)

p_mean_exp2 <- two_parts_df_exp2 %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  mutate(
    num_chosen = case_when(
      response_position == "middle" ~ middle_number,
      response_position == "left" ~ left_number,
      response_position == "right" ~ right_number
    )
  ) %>%
  summarise(mean_num_chosen = mean(num_chosen)) %>%
  mutate(diff = mean_num_chosen - 3)

p_mean_exp3 <- two_parts_df_exp3 %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  mutate(
    num_chosen = case_when(
      response_position == "middle" ~ middle_number,
      response_position == "left" ~ left_number,
      response_position == "right" ~ right_number
    )
  ) %>%
  summarise(mean_num_chosen = mean(num_chosen)) %>%
  mutate(diff = mean_num_chosen - 3)

```


ppts in exp. 1, mean chosen number `r p_mean_exp1 %>% pull(mean_num_chosen) %>% mean() %>% printnum()`. t-test against 3, `r apa_print(t.test(p_mean_exp2$diff))$statistic`

ppts in exp. 2, mean chosen number `r p_mean_exp2 %>% pull(mean_num_chosen) %>% mean() %>% printnum()`. t-test against 3, `r apa_print(t.test(p_mean_exp2$diff))$statistic`

ppts in exp. 3, mean chosen number `r p_mean_exp3 %>% pull(mean_num_chosen) %>% mean() %>% printnum()`. t-test against 3, `r apa_print(t.test(p_mean_exp2$diff))$statistic`



given incorrect trials in part 2, P(higher_number) in trials where participant didnt chose only stones in part 1 and reconstructed incorrectly. estimated exp. 1 p(higher_number)  = `r plogis(round(fixef(model_exp1)[1], 2))` (β = `r printnum(round(fixef(model_exp1)[1], 2))`, 
SE = `r printnum(round(summary(model_exp1)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(model_exp1)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(model_exp1)$coefficients[1, "Pr(>|z|)"])`)

In exp. 2 p(higher_number) = = `r plogis(round(fixef(model_exp2)[1], 2))` (β = `r printnum(round(fixef(model_exp2)[1], 2))`, 
SE = `r printnum(round(summary(model_exp2)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(model_exp2)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(model_exp2)$coefficients[1, "Pr(>|z|)"])`)





```{r pstones_diff, cache=F}

exp1_pstones_diff <- two_parts_df_exp1 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id) %>%
  summarise(
    pstone_p1 = mean(chose_onlystones),
    pstone_p2 = mean(chose_onlystones_p2)
  ) %>%
  mutate(
    pstone_diff = pstone_p2 - pstone_p1
  )

t.test(exp1_pstones_diff$pstone_diff)


exp2_pstones_diff <- two_parts_df_exp2 %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id) %>%
  summarise(
    pstone_p1 = mean(chose_onlystones),
    pstone_p2 = mean(chose_onlystones_p2)
  ) %>%
  mutate(
    pstone_diff = pstone_p2 - pstone_p1
  )

t.test(exp2_pstones_diff$pstone_diff)




```


Exp. 1 within-subject t-test, pstones part 2 - pstones part 1 `r apa_print(t.test(exp1_pstones_diff$pstone_diff))$statistic`

Exp. 2 within-subject t-test, pstones part 2 - pstones part 1 `r apa_print(t.test(exp2_pstones_diff$pstone_diff))$statistic`



```{r estimate_means, fig.width = 8, fig.height = 3}

library(patchwork) 


exp2_chose_gems <- two_parts_df_exp2 %>%
  group_by(subj_id) %>% 
  summarise(pgem = mean(chose_gems))

exp2_estimate <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(pgem = first(number_correct_part1)/50)

exp2_chose_gems_p2 <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(pgem = mean(chose_gems_p2))

library(dplyr)
library(ggplot2)
library(ggtext) 


combined <- bind_rows(
  exp2_chose_gems %>% mutate(exp = "Part 1"),
  exp2_chose_gems_p2 %>% mutate(exp = "Part 2"),
  exp2_estimate %>% mutate(exp = "Part 1 Estimate")
) %>%
  mutate(exp = factor(exp, levels = c("Part 1", "Part 2", "Part 1 Estimate")))
  

summary_df <- combined %>%
  group_by(exp) %>%
  summarise(
    mean_pgem = mean(pgem, na.rm = TRUE),
    se_pgem   = sd(pgem, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )



asterisk_df_pstones <- summary_df %>%
  mutate(
    label = "***",
    y = mean_pgem + se_pgem + 0.01
  )




exp2_estimate <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(pgem = first(number_correct_part1)/50)

exp2_chose_gems <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(pgem = mean(chose_gems, na.rm = TRUE), .groups = "drop")

exp2_chose_gems_p2 <- two_parts_df_exp2 %>%
  group_by(subj_id) %>%
  summarise(pgem = mean(chose_gems_p2, na.rm = TRUE), .groups = "drop")

p1_vs_estimate <- exp2_estimate %>%
  mutate(pgem_estimate = pgem) %>%
  select(subj_id, pgem_estimate) %>%
  inner_join(
    exp2_chose_gems,
    by = "subj_id"
  )

p2_vs_estimate <- exp2_estimate %>%
  mutate(pgem_estimate = pgem) %>%
  select(subj_id, pgem_estimate) %>%
  inner_join(
    exp2_chose_gems_p2,
    by = "subj_id"
  )

p1_vs_p2 <- exp2_chose_gems_p2 %>%
  mutate(pgem_p2 = pgem) %>%
  select(subj_id, pgem_p2) %>%
  inner_join(
    exp2_chose_gems,
    by = "subj_id"
  )

fmt_p <- function(p) {
  if (is.na(p)) return(NA_character_)
  if (p < 0.001) return("***")
  if (p < 0.01) return ("**")
  if (p < 0.05) return ("*")
  sprintf("p = %.2f", p)
}

p1_p2_ttest <- fmt_p(t.test(p1_vs_p2$pgem_p2, p1_vs_p2$pgem, paired = TRUE)$p.value)
p1_estimate <- fmt_p(t.test(p1_vs_estimate$pgem_estimate, p1_vs_estimate$pgem, paired = TRUE)$p.value)
p2_estimate <- fmt_p(t.test(p2_vs_estimate$pgem_estimate, p2_vs_estimate$pgem, paired = TRUE)$p.value)


x_p1       <- which(levels(summary_df$exp) == "Part 1")
x_p2       <- which(levels(summary_df$exp) == "Part 2")
x_estimate <- which(levels(summary_df$exp) == "Part 1 Estimate")


y_top <- max(summary_df$mean_pgem + summary_df$se_pgem, na.rm = TRUE)
pad   <- diff(range(c(0.35, 0.45))) * 0.06
y_p1_p2  <- 0.365   
y_p1_est <- 0.4375  
y_p2_est <- 0.355   

cap_height <- 0.005
line_w     <- 0.6
text_pad   <- 0.006
text_size  <- 5

seg_df_pgem <- tibble::tibble(
  x       = c(x_p1, x_p1,  x_p2),
  xend    = c(x_p2, x_estimate, x_estimate),
  y       = c(y_p1_p2, y_p1_est, y_p2_est),
  label   = c(p1_p2_ttest, p1_estimate, p2_estimate),
  cap_dir = c(1, -1, 1)
)
seg_up <- seg_df_pgem %>%
  dplyr::filter(cap_dir == 1) %>%
  dplyr::mutate(yend = y + cap_height)

seg_down <- seg_df_pgem %>%
  dplyr::filter(cap_dir == -1) %>%
  dplyr::mutate(yend = y - cap_height)


text_df_pgem <- seg_df_pgem %>%
  dplyr::mutate(
    x = (x + xend) / 2,
    y = ifelse(cap_dir == -1, y + text_pad, y - text_pad),
    fontface = ifelse(label %in% c("*", "**", "***"), "bold", "plain")
  ) %>%
  dplyr::select(x, y, label, fontface)

pgem_estimate <- ggplot(summary_df, aes(x = exp, y = mean_pgem)) +
  geom_errorbar(
    aes(ymin = mean_pgem - se_pgem,
        ymax = mean_pgem + se_pgem),
    width = 0.12,
    linewidth = 0.8
  ) +
  geom_point(
    shape = 23, size = 3, fill = "black", color = "black"
  ) +
  scale_y_continuous(
    limits = c(0.33, 0.45),
    breaks = seq(0.1, 1.1, by = 0.1),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(x = "", y = "P(*gems*)") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_markdown(size = 12, hjust = 0.5),
        legend.position = "none",
        axis.title.y = element_markdown(size = 14),
        axis.title.x = element_markdown(size = 14),
        axis.text = element_text(size = 12))


pgem_estimate <- pgem_estimate +
  geom_segment(
    data = seg_df_pgem,
    mapping = aes(x = x, xend = xend, y = y, yend = y),
    inherit.aes = FALSE,
    color = "black",
    linewidth = line_w,
    lineend = "butt"
  )

if (nrow(seg_up) > 0) {
  pgem_estimate <- pgem_estimate +
    geom_segment(
      data = seg_up,
      mapping = aes(x = x, xend = x, y = y, yend = yend),
      inherit.aes = FALSE,
      color = "black",
      linewidth = line_w
    ) +
    geom_segment(
      data = seg_up,
      mapping = aes(x = xend, xend = xend, y = y, yend = yend),
      inherit.aes = FALSE,
      color = "black",
      linewidth = line_w
    )
}

if (nrow(seg_down) > 0) {
  pgem_estimate <- pgem_estimate +
    geom_segment(
      data = seg_down,
      mapping = aes(x = x, xend = x, y = y, yend = yend),
      inherit.aes = FALSE,
      color = "black",
      linewidth = line_w
    ) +
    geom_segment(
      data = seg_down,
      mapping = aes(x = xend, xend = xend, y = y, yend = yend),
      inherit.aes = FALSE,
      color = "black",
      linewidth = line_w
    )
}

pgem_estimate <- pgem_estimate +
  geom_text(
    data = text_df_pgem,
    mapping = aes(x = x, y = y, label = label, fontface = fontface),
    inherit.aes = FALSE,
    size = text_size,
    vjust = 0.5
  )


print(pgem_estimate)

ggsave(
  "pgem_estimate.png",
  pgem_estimate,
  width = 8,
  height = 4,
  dpi = 300
)

```

Difference in means for part 1 p(gems) and part 2 p(gems): `r apa_print(t.test(p1_vs_p2$pgem_p2, p1_vs_p2$pgem, paired = TRUE))$statistic`

Difference in means for part 1 p(gems) and part 1 p(gems) estimate: `r apa_print(t.test(p1_vs_estimate$pgem_estimate, p1_vs_estimate$pgem, paired = TRUE))$statistic`

Difference in means for part 1 p(gems) estimate and part 2 p(gems): `r apa_print(t.test(p2_vs_estimate$pgem_estimate, p2_vs_estimate$pgem, paired = TRUE))$statistic`

```{r conf_analysis, cache=F}

correct_conf_summary <- two_parts_df_exp2 %>%
  filter(!is_catch_trial, is_consistent) %>%
  group_by(subj_id) %>%
  summarise(mean_conf_subj = mean(confidence, na.rm = TRUE)) %>%
  summarise(
    mean_conf = mean(mean_conf_subj),
    se_conf = sd(mean_conf_subj)/sqrt(n()),
    n = n()
  ) 

incorrect_conf_summary <- two_parts_df_exp2 %>%
  filter(!is_catch_trial, !is_consistent) %>%
  group_by(subj_id) %>%
  summarise(mean_conf_subj = mean(confidence, na.rm = TRUE)) %>%
  summarise(
    mean_conf = mean(mean_conf_subj),
    se_conf = sd(mean_conf_subj)/sqrt(n()),
    n = n()
  ) 


gems_stones_conf <- two_parts_df_exp2 %>%
  filter(!is_catch_trial, chose_gems, !is_consistent)%>%
  group_by(subj_id) %>%
  summarise(mean_conf_subj = mean(confidence)) %>%
  summarise(
    mean_conf = mean(mean_conf_subj),
    se_conf = sd(mean_conf_subj)/sqrt(n()),
    n = n()
  )

stones_gems_conf <- two_parts_df_exp2 %>% 
  filter(!is_catch_trial, !chose_gems, chose_gems_p2) %>%
  group_by(subj_id) %>%
  summarise(mean_conf_subj = mean(confidence)) %>%
  summarise(
    mean_conf = mean(mean_conf_subj),
    se_conf = sd(mean_conf_subj)/sqrt(n()),
    n = n()
  )

stones_stones_conf <- two_parts_df_exp2 %>%
  filter(!is_catch_trial, !chose_gems, !chose_gems_p2, !is_consistent) %>%
  group_by(subj_id) %>%
  summarise(mean_conf_subj = mean(confidence)) %>%
  summarise(
    mean_conf = mean(mean_conf_subj),
    se_conf = sd(mean_conf_subj)/sqrt(n()),
    n = n()
  )



```

(all part 2 incorrect trials)
gems -> stones mean confidence `r printnum(gems_stones_conf$mean_conf)`, SE = `r printnum(gems_stones_conf$se_conf)`

stones -> gems mean confidence `r printnum(stones_gems_conf$mean_conf)`, SE = `r printnum(stones_gems_conf$se_conf)`

stones -> stones mean confidence `r printnum(stones_stones_conf$mean_conf)`, SE = `r printnum(stones_stones_conf$se_conf)`

# Results

# Discussion


\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
